---
title: "1.2 Summary Statistics"
author: "Paul Adamson"
date: "November 25, 2015"
output: html_document
---
In Example 1.9, we compute the mean of a random sample of the heights of five men.
```{r}
exm1_9<-read.csv("./ascii_datasets/ascii-comma/Ch1/exm1-9.txt")
mean(exm1_9$Heights)
```
In Example 1.10, we compute the variance and standard deviation of the data used in 
Example 1.9.
```{r}
var(exm1_9$Heights)
sd(exm1_9$Heights)
```
In Example 1.11, we find the sample median of the data used in Example 1.9.
```{r}
median(exm1_9$Heights)
```

In Example 1.12, we compute the mean, median, and the 5%, 10%, and 20% trimmed means of
some provided data.
```{r}
exm1_12<-read.csv("./ascii_datasets/ascii-comma/Ch1/exm1-12.txt")
mean(exm1_12$Stress)
median(exm1_12$Stress)
mean(exm1_12$Stress, trim=0.05)
mean(exm1_12$Stress, trim=0.1)
mean(exm1_12$Stress, trim=0.2)
```
Note that the `trim` argument behaves slightly differently than the way that the 
"p% trimmed mean" is defined in the text. the `mean` function always rounds the number
of trimmed values down. Therefore, the 20% trimmed mean as computed with the `mean`
function in R is equivalent to the mean of the 5th through 20th values.
```{r}
mean(exm1_12$Stress[5:20])
```
In Example 1.13, we find the mode and range of the sample from Example 1.12. R does
not include a mode function, so we must define one (by adapting 
[Chris' solution on stackoverflow](http://stackoverflow.com/questions/2547402/standard-library-function-in-r-for-finding-the-mode)). The following function `modeav` takes the
following options for the `method` argument:

- method = "mode" [default]: calculates the mode for a unimodal vector, else returns an NA
- method = "nmodes": calculates the number of modes in the vector
- method = "modes": lists all the modes for a unimodal or polymodal vector

```{r}
modeav <- function (x, method = "mode", na.rm = FALSE)
{
  x <- unlist(x)
  if (na.rm)
    x <- x[!is.na(x)]
  u <- unique(x)
  n <- length(u)
  #get frequencies of each of the unique values in the vector
  frequencies <- rep(0, n)
  for (i in seq_len(n)) {
    if (is.na(u[i])) {
      frequencies[i] <- sum(is.na(x))
    }
    else {
      frequencies[i] <- sum(x == u[i], na.rm = TRUE)
    }
  }
  #mode if a unimodal vector, else NA
  if (method == "mode" | is.na(method) | method == "")
  {return(ifelse(length(frequencies[frequencies==max(frequencies)])>1,NA,u[which.max(frequencies)]))}
  #number of modes
  if(method == "nmode" | method == "nmodes")
  {return(length(frequencies[frequencies==max(frequencies)]))}
  #list of all modes
  if (method == "modes" | method == "modevalues")
  {return(u[which(frequencies==max(frequencies), arr.ind = FALSE, useNames = FALSE)])}  
  #error trap the method
  warning("Warning: method not recognised.  Valid methods are 'mode' [default], 'nmodes' and 'modes'")
  return()
}
```
And the solution is now straightforward.
```{r}
modeav(exm1_12$Stress, method="modes")
rng<-range(exm1_12$Stress)
rng[2]-rng[1]
```
In Example 1.14, the first and third quartiles are computed.
```{r}
quantile(exm1_12$Stress, probs=c(0.25, 0.75), type=2)
```
In Example 1.15, the 65th percentile is computed.
```{r}
quantile(exm1_12$Stress,probs=c(0.65), type=2)
```
Figure 1.4 provides some descriptive statistics of particulate matter emissions
data at two different altitudes.
```{r}
table1_1<-read.csv("./ascii_datasets/ascii-comma/Ch1/table1-1.txt")
table1_2<-read.csv("./ascii_datasets/ascii-comma/Ch1/table1-2.txt")
library(plyr)
summarize(table1_1,N=length(PM),sd=sd(PM))
summary(table1_1$PM)
summarize(table1_2,N=length(PM),sd=sd(PM))
summary(table1_2$PM)
```
In Example 1.16, frequencies and sample proportions are computed from summary
information from categorical data.
```{r}
exm1_16<-read.csv("./ascii_datasets/ascii-comma/Ch1/exm1-16.txt")
freq<-exm1_16$Number
freq
sampleproportions<-exm1_16$Number/1000
sampleproportions
```