---
title: '1.3 Graphical Summaries'
author: "Paul Adamson"
date: "November 26, 2015"
output: html_document
---
Stem-and-leaf plots are a type of graphical summary of data used to visualize lists.
```{r}
table1_3<-read.csv("./ascii_datasets/ascii-comma/Ch1/table1-3.txt")
stem(table1_3$Duration)
stem(table2_9$correct,scale=2)
table1_2<-read.csv("./ascii_datasets/ascii-comma/Ch1/table1-2.txt")
stem(table1_2$PM)
```
A dotplot is a graph that can give the rough impression of the shape of some data.
```{r}
stripchart(table1_3$Duration,
           method="stack",
           pch=19,
           col="Red",
           offset=0.5,
           main="Old Faithful geyser dormant period duration (min)",
           xlab="Time (min)",
           xlim=c(40,100),
           frame.plot = FALSE)
```
```{r}
minClassValue<-1
maxClassValue<-25
widthOfClasses<-2
breaks = seq(minClassValue,maxClassValue,by=widthOfClasses)
table1_2.cut = cut(table1_2$PM,breaks, right=FALSE)
table1_4<-as.data.frame(table(table1_2.cut))
names(table1_4)<-c("interval","frequency")
table1_4
table1_4<-mutate(table1_4,
                 relativefrequency=frequency/sum(table1_4$frequency),
                 density=relativefrequency/widthOfClasses)
table1_4
barplot(table1_4$relativefrequency,
        names.arg=table1_4$interval, 
        axis.lty=0, 
        xlab="Emissions (g/gal)", 
        ylab="Relative Frequency", 
        ylim=c(0,.3),
        plot=TRUE,
        space=c(0))
```

```{r}
minClassValue<-1
maxClassValue<-11
widthOfClasses<-2
breaks = seq(minClassValue,maxClassValue,by=widthOfClasses)
breaks<-c(breaks,15,25)
table1_2.cut = cut(table1_2$PM,breaks, right=FALSE)
table1_5<-as.data.frame(table(table1_2.cut))
names(table1_5)<-c("interval","frequency")
width<-numeric(0)
for (i in 1:nrow(table1_5)) {
 width[i] <- breaks[i+1] - breaks[i]
}
table1_5<-cbind(table1_5,width)
table1_5<-mutate(table1_5,
                 relativefrequency=frequency/sum(table1_5$frequency),
                 density=relativefrequency/width)
table1_5
barplot(table1_5$density,table1_5$width,
        names.arg=table1_5$interval, 
        axis.lty=0, 
        xlab="Emissions (g/gal)", 
        ylab="Density", 
        ylim=c(0,.15),
        plot=TRUE,
        space=c(0))
