---
title: "2.1 Basic Ideas"
author: "Paul Adamson"
date: "November 27, 2015"
output: html_document
---
Example 2.1
```{r}
require(dplyr)
a<-c(9,10,11,12)
b<-c(18,19,20,21)
S<-arrange(expand.grid(a=a,b=b),a)
A<-filter(S,a>10)
B<-filter(S,b<19)
C<-filter(S,a+b==28)
S
A
B
C
```
Example 2.2
```{r}
arrange(union(B,C),b,a)
arrange(intersect(A,anti_join(S,B)),a,b)
```
For Example 2.3, we can determine if two events are mutually exclusive by finding
their intersection. By this method we see that A and C are mutually exclusive.
```{r}
intersect(A,B)
intersect(B,C)
intersect(A,C)
```
For Example 2.4, we can build a data frame with a single row, where each column
is the probability for one of the mutually exclusive events.
```{r}
P<-data.frame("bullseye"=0.1,"innerring"=.25,"outerring"=0.45)
```
Then, we can use the `mutate` function from the `dplyr` package to add columns
for the "hits target" and "misses target" probabilities.
```{r}
P<-mutate(P,hitstarget=P$bullseye+P$innerring+P$outerring)
P$hitstarget
P<-mutate(P,missestarget=1-P$hitstarget)
P$missestarget
```
Example 2.5
```{r}
example2_5<-read.csv("./ascii_datasets/ascii-comma/Ch2/exm2-5.txt")
S<-data.frame(Number.of.Crashes=c(0,1,2,3,4))
S
A<-filter(S,Number.of.Crashes>2)
A
B<-filter(S,Number.of.Crashes>=1)
B
A.indices.in.S<-na.omit(match(A$Number.of.Crashes,S$Number.of.Crashes))
sum(example2_5$Probability[A.indices.in.S])
Bc.indices.in.S<-na.omit(match(anti_join(S,B)$Number.of.Crashes,S$Number.of.Crashes))
1-sum(example2_5$Probability[Bc.indices.in.S])
B.indices.in.S<-na.omit(match(B$Number.of.Crashes,S$Number.of.Crashes))
sum(example2_5$Probability[B.indices.in.S])
```
Example 2.6
```{r}
example2_6<-read.csv("./ascii_datasets/ascii-comma/Ch2/exm2-6.txt")
names(example2_6)<-c("length","toothin","correctdiameter","toothick")
example2_6$length<-c("tooshort","correctlength","toolong")
require(data.table)
example2_6.melted<-melt(example2_6)
names(example2_6.melted)<-c("length","diameter","frequency")
tooshort<-sum(filter(example2_6.melted,length=="tooshort")$frequency)
total<-sum(example2_6.melted$frequency)
tooshort/total
```
Example 2.7
```{r}
tooshort.or.toothick<-sum(filter(example2_6.melted,length=="tooshort" | diameter=="toothick")$frequency)
tooshort.or.toothick/total
```
