---
title: "2.3 Conditional Probability and Independence"
author: "Paul Adamson"
date: "November 28, 2015"
output: html_document
---
```{r}
example2_6<-read.csv("./ascii_datasets/ascii-comma/Ch2/exm2-6.txt")
names(example2_6)<-c("length","toothin","correctdiameter","toothick")
example2_6$length<-c("tooshort","correctlength","toolong")
require(dplyr)
require(data.table)
example2_6.melted<-melt(example2_6)
names(example2_6.melted)<-c("length","diameter","frequency")
example2_6.melted
correctdiameter<-sum(filter(example2_6.melted,diameter=="correctdiameter")$frequency)
total<-sum(example2_6.melted$frequency)
correctdiameter/total
correctlength<-sum(filter(example2_6.melted,length=="correctlength")$frequency)
correctdiameter.given.correctlength<-sum(filter(example2_6.melted,length=="correctlength",
                                                diameter=="correctdiameter")$frequency)
correctdiameter.given.correctlength/correctlength
```

Example 2.17
```{r}
correctdiameter.given.toolong<-sum(filter(example2_6.melted,length=="toolong",
                                                diameter=="correctdiameter")$frequency)
toolong<-sum(filter(example2_6.melted,length=="toolong")$frequency)
correctdiameter.given.toolong/toolong
correctdiameter/total
```

Example 2.18
```{r}
0.01/0.03
```

Example 2.19
```{r}
0.01/0.02
```

Example 2.20
```{r}
toolong.given.toothin<-sum(filter(example2_6.melted,length=="toolong",
                                                diameter=="toothin")$frequency)
toolong<-sum(filter(example2_6.melted,length=="toolong")$frequency)
toothin<-sum(filter(example2_6.melted,diameter=="toothin")$frequency)
toolong/total
toolong.given.toothin/toothin
```
Example 2.21
```{r}
0.05*0.10
```

Example 2.22
```{r}
(1-0.08)*(1-0.05)
```

Example 2.23
```{r}
(1-0.20)**5
```

Example 2.24
```{r}
1-0.20**5
```

Example 2.25
```{r}
require(dplyr)
carengines<-data.frame(enginesize=c("small","medium","large"),density=c(.45,.35,.20),
                       emissionsfail=c(.1,.12,.15))
sum(carengines$density*carengines$emissionsfail)
```

Example 2.26
```{r}
(0.99*0.005)/(0.99*0.005+0.01*0.995)
```

Example 2.27
```{r}
carengines<-mutate(carengines,fail.given.enginesize=density*emissionsfail)
filter(carengines,enginesize=="small")$fail.given.enginesize/sum(carengines$fail.given.enginesize)
```

Example 2.28
```{r}
0.98*0.95
```

Example 2.29
```{r}
0.90+0.85-0.90*0.85
```

Example 2.30 is not presented here yet. It might be nice to develop a package that
computes system reliability given a matrix that describes the connections and
component reliabilities. The package could also use the `diagram` package to present
a flow diagram.


