---
title: "3.1 Measures of Location"
author: "Paul Adamson"
date: "November 22, 2015"
output: html_document
---
Mean
====
Begin by loading the `xlsx` package,

```{r}
library(xlsx)
```
and then the monthly starting salary data:
```{r}
table3_1<-read.xlsx("../../anderson9edata/Excel Files/Ch 03 Descriptive/Salary.xls",1)
names(table3_1)<-c("graduate","monthlystartingsalary")
table3_1
```

The mean monthly salary is computed with one line:

```{r}
mean(table3_1$monthlystartingsalary)
```

Median
======
```{r}
median(table3_1$monthlystartingsalary)
```
Mode
====
R does not have a built-in mode function. (Well, actually the built-in `mode` function
is used to get or set the type or storage mode of an object, not to calculate the statistical
mode or *modus* of a collection of values.) Therefore, we define a function `modus` that should
work for a vector of numeric, character, or factor data.
```{r}
modus <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```
We then read in the soft drink data from Table 2.2 and compute the modus.
```{r}
softdrink<-read.xlsx("../../anderson9edata/Excel Files/Ch 02 Descriptive/SoftDrink.xls",1)
modus(softdrink$Brand.Purchased)
```
Percentiles
===========
Anderson uses averaging at discontinuities, or the *Type 2* algorithm of the `quantile` function
in R.
```{r}
quantile(table3_1$monthlystartingsalary,probs=c(.85,.50),type=2)
```
Quartiles
=========
Division into quartiles is the default for the `quantile` function in R.
```{r}
quantile(table3_1$monthlystartingsalary,type=2)
```