---
title: "3.2 Measures of Variability"
author: "Paul Adamson"
date: "November 22, 2015"
output: html_document
---
Range
=====
Begin by loading the `xlsx` package,
```{r}
library(xlsx)
```
and then the monthly starting salary data,
```{r}
table3_1<-read.xlsx("../../anderson9edata/Excel Files/Ch 03 Descriptive/Salary.xls",1)
names(table3_1)<-c("graduate","monthlystartingsalary")
```
and compute the range
```{r}
range(table3_1$monthlystartingsalary)
```
Interquartile Range
===================
Just as in section 3.1, we will use Type 2 algorithm of the `quantile` function that is
called by the `IQR` function to compute the interquartile range.
```{r}
IQR(table3_1$monthlystartingsalary,type=2)
```
Variance
========
Class size data as given in Section 3.1 is input.

```{r}
library(dplyr)
table3_3<-data.frame(numberofstudentsinclass=c(46,54,42,46,32))
```
Use mutate from `dplyr` package to compute the mean, deviation about the mean, and squared 
deviation about the mean.
```{r}
table3_3<-mutate(table3_3,
                  mean=mean(numberofstudentsinclass),
                  deviationaboutmean=numberofstudentsinclass-mean,
                  squareddeviationaboutmean=deviationaboutmean**2)
```
And again for the data in Table 3.1. Note this time, we do not include the column of mean
values. The mean of all of the observations of a variable is not an observation and should
not be repeated on every row of a table. It is computed and stored in a separate vector
`samplemean` to be used in the `mutate` function.
```{r}
table3_4<-data.frame(table3_1$monthlystartingsalary)
names(table3_4)<-c("monthlysalary")
samplemean<-mean(table3_4$monthlysalary)
table3_4<-mutate(table3_4,
                  deviationaboutmean=monthlysalary-samplemean,
                  squareddeviationaboutmean=deviationaboutmean**2)
table3_4
```

Standard Deviation
==================
Standard deviation for the sample of class sizes:
```{r}
sd(table3_3$numberofstudentsinclass)
```
Standard deviation for the data on starting salaries:
```{r}
sd(table3_4$monthlysalary)
```

Coefficient of Variation
========================
The `cv` function defined below computes the coefficient of variation as defined in Equation 3.8.
```{r}
cv <- function(x){
  samplesd<-sd(x)
  samplemean<-mean(x)
  (samplesd/samplemean)*100
}

cv(table3_3$numberofstudentsinclass)
cv(table3_4$monthlysalary)
```